The Sound of Teaching Music: Expert Pianists' Performance Modulations for Novices

### 2020-12-16 paper/manuscript.tex
modified discussion/correct some stats report
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

\abstract{
aaaa
}

@@

Importantly, in some domains teaching requires not only demonstrating \emph{what} to do, but \emph{how} to perform actions. In artistic contexts, novices need to learn how exactly to perform actions by relying on particular techniques (Sloboda, 2000). For example, when playing a piece of music on the piano, it is not sufficient to press the keys in the correct order, but learners need to be able to implement expressive techniques (Juslin, 2003). This raises the question of whether and how experts modulate their actions during demonstration in artistic contexts, where expressivity is an integral part of what learners need to acquire. Studying action modulations that serve the teaching of artistic techniques is of broader theoretical importance: It can show whether action modulations can be used to selectively highlight subtle aspects of actions that define expert performance.

@@

If experts rely on generic action modulations, it can be expected that they will play more slowly during teaching demonstration, regardless of the kind of expressive techniques to be taught. To the extent that experts use action modulations to support their teaching of specific expressive techniques, they should produce longer legato and shorter staccato when teaching articulation whereas they should produce louder forte and softer piano when teaching dynamics. Furthermore, one could speculate that they might produce modulations specifically at structurally important points that best highlight the technique to be taught. Importantly, they should avoid modulating irrelevant properties of expression, e.g., the smoothness of the sound while teaching dynamics.

In Experiment 1, we employed a simple musical scale to examine whether and how skilled pianists varied depending on which expressive techniques (i.e., either articulation or dynamics) was taught. Experiment 2 was conducted to replicate our findings from Experiment 1 with a more realistic piece of music.

@@

In order to compare the mean IOIs between the teaching and performing condition, we conducted a Wiscoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution. The Wiscoxon Signed-rank test revealed that participants played more slowly in the teaching condition {[}\emph{Mdn} = 189.55 (ms), \emph{IQR} = 26.74{]} than in the performing condition {[}\emph{Mdn} = 185.24 (ms), \emph{IQR} = 19.32{]} while playing the piece with notated articulation (\emph{p} = 0.021; \emph{Figure \ref{fig:ioi-1}}).

@@

A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) revealed that there was a significant main effect of Subcomponent-LS (\emph{F}(1,30) = 1158, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.94) and a significant interaction between Condition and Subcomponent-LS (\emph{F}(1,30) = 8.31, \emph{p} = 0.007, \(\eta_G^2\) = 0.01). Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that participants produced shorter staccato in the teaching condition than in the performing condition (\emph{p} = 0.004) while there was no significant difference in legato between the teaching and performing condition (\emph{p} = 0.32).

@@

A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) showed that there was a significant interaction between Condition and Subcomponent-LS (\emph{F}(1,30) = 7.38, \emph{p} = 0.011, \(\eta_G^2\) = 0.0044). Post-hoc comparisons based on the estimated marginal means with Tukey adjustment did not show a significant difference between the teaching and performing condition for each subcomponent (Legato; \emph{p} = 0.696, Staccato; \emph{p} = 0.07).

@@

The Wiscoxon Signed-rank test revealed no significant difference between the teaching condition {[}\emph{Mdn} = 186.29 (ms), \emph{IQR} = 21.83{]} and the performing condition {[}\emph{Mdn} = 186.27 (ms), \emph{IQR} = 19.58{]} while playing the piece with notated dynamics (\emph{p} = 0.111; \emph{Figure \ref{fig:ioi-1}}).

@@

A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,30) = 369.54, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.7367) and a significant interaction between Condition and Subcomponent-FP (\emph{F}(1,30) = 6.80, \emph{p} = 0.014, \(\eta_G^2\) = 0.0034). Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte in the teaching condition than in the performing condition (\emph{p} = 0.021) while there was no significant difference in piano between the teaching and performing condition (\emph{p} = 0.448).

@@

Additionally, we conducted an exploratory analysis by focusing on specific points at which each subcomponent changed from one to the other (i.e., forte to piano; FtoP, piano to forte; PtoF). These points could be structurally important in order to make a contrast between forte and piano. We calculated a KV difference for each interval subtracting the KV value of the current note from that of the following note. A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (FtoP vs.~PtoF) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,30) = 223.89, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.8344) and a significant interaction of Condition and Subcomponent-FP (\emph{F}(1,30) = 25.74, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.0321). Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano and from piano to forte in the teaching condition than in the performing condition (FtoP; \emph{p} = 0.002, PtoF; \emph{p} \textless{} 0.001.

\begin{verbatim}
## Warning: Removed 2 rows containing missing values (geom_text).
\end{verbatim}

@@

A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) revealed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,30) = 377, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.31), reflecting more key overlap for forte notes compared to piano notes. Neither the main effect of Condition (\emph{F}(1,30) = 0.423, \emph{p} = 0.520, \(\eta_G^2\) = 0.00) nor the interaction between Condition and Subcomponent-FP (\emph{F}(1,30) = 0.000, \emph{p} = 0.987, \(\eta_G^2\) = 0.00).

@@

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-vel-diff-1-1} \caption{\label{fig:vel-diff-1}Experiment 1: KV Difference at transition points when playing the piece with dynamics. Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box. Significance levels: * \textit{p} < .05, ** \textit{p} < .01, *** \textit{p} < .001}\label{fig:plot-vel-diff-1}
\end{figure}

The findings from Experiment 1 indicated that skilled pianists were able to modify their performance for teaching purposes. For IOIs (tempo), although there was a general trend of slowing down, we found a small but significant difference only when playing the piece with the notated articulation. This could be due to the leading metronome, which was given prior to each performance. As participants tried to follow the ideal tempo through their performances, it was possible that there was only a tiny difference between the two conditions. Another possibility could be that slower performance may be more or only beneficial to highlight the relation between two notes (i.e.~to what extent two notes overlap) when teaching articulation whereas such relation was irrelevant when teaching dynamics. With regard to KOT (articulation) and KV (dynamics), participants successfully highlight the relevant aspect of the technique. Specifically, they produced shorter staccato when teaching articulation and produced louder forte when teaching dynamics. Moreover, our exploratory analysis demonstrated that participants made a larger contrast in dynamics at transition points where one of two subcomponents moved to the other (i.e., forte to piano or piano to forte). These findings confirmed that participants modulated their performance in systematic and fine-grained ways in order to teach expressive techniques.

One of the limitations of Experiment 1 is that we employed a simple musical scale, which was good to gain an experimental control but not a realistic and natural stimulus to implement artistic expressions. In Experiment 2, we chose one piece of music from actual piano pieces and modified it for the purpose of the experiment. We predict that participants will played more slowly regardless of the techniques to be taught or only when teaching articulation based on the finding from Experiment 1. Also, we expect that participants will be able to highlight only the relevant aspects of the techniques to be taught. More precisely, they should be able to produce longer legato and shorter staccato while teaching articulation. When teaching dynamics, they should be able to produce louder forte and softer piano. Given the findings we observed in Experiment 1, participants will make a larger contrast between forte and piano at transition points.

@@

A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) showed that there was a significant main effect of Subcomponent-LS (\emph{F}(1,18) = 2324, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.98) and a significant interaction between Condition and Subcomponent-LS (\emph{F}(1,18) = 6.90, \emph{p} = 0.017, \(\eta_G^2\) = 0.01). Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that participants produced shorter staccato in the teaching condition than in the performing condition (\emph{p} = 0.001) while there was no significant difference in legato between the teaching and performing condition (\emph{p} = 0.85).

@@

A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) showed that there was a significant main effect of Subcomponent-LS (\emph{F}(1,18) = 8.15, \emph{p} = 0.011, \(\eta_G^2\) = 0.0400), reflecting louder sound for legato notes compared to staccato notes. Neither the main effect of Condition (\emph{F}(1,18) = 0.46, \emph{p} = 0.506, \(\eta_G^2\) = 0.0008) nor the interaction between Condition and Subcomponent-LS (\emph{F}(1,18) = 8.15, \emph{p} = 0.799, \(\eta_G^2\) = 0.0001).

@@

A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,18) = 126.10, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.7190) and a significant interaction between Condition and Subcomponent-FP (\emph{F}(1,18) = 6.11, \emph{p} = 0.024, \(\eta_G^2\) = 0.0217). Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte and softer piano in the teaching condition than in the performing condition (Forte; \emph{p} = 0.036, Piano; \emph{p} = 0.028).

@@

A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (FtoP vs.~PtoF) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,18) = 110.64, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.8167) and a significant interaction of Condition and Subcomponent-FP (\emph{F}(1,18) = 9.77, \emph{p} = 0.006, \(\eta_G^2\) = 0.0641). Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano and from piano to forte in the teaching condition than in the performing condition (FtoP; \emph{p} = 0.017, PtoF; \emph{p} = 0.003.

\begin{verbatim}
## Warning: Removed 2 rows containing missing values (geom_text).
\end{verbatim}

@@

A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,18) = 559, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.93), reflecting more key overlap for forte notes compared to piano notes. Neither the main effect of Condition (\emph{F}(1,18) = 0.016, \emph{p} = 0.902, \(\eta_G^2\) = 0.00) nor the interaction between Condition and Subcomponent-FP (\emph{F}(1,18) = 0.576, \emph{p} = 0.458, \(\eta_G^2\) = 0.00).

@@

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-kot-2-1} \caption{\label{fig:kot-2}Experiment 2: KOR when playing the piece with either articulation (left) or dynamics (right). Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box. Significance levels: * \textit{p} < .05, ** \textit{p} < .01, *** \textit{p} < .001}\label{fig:plot-kot-2}
\end{figure}

@@

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-vel-diff-2-1} \caption{\label{fig:vel-diff-2}Experiment 2: KV Difference at transition points when playing the piece with dynamics. Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box. Significance levels: * \textit{p} < .05, ** \textit{p} < .01, *** \textit{p} < .001}\label{fig:plot-vel-diff-2}
\end{figure}

Blah

@@

@@

@@

@@

@@

\endgroup
\raggedbottom

>>>>

\abstract{
Experts modulate their performance of actions for teaching purposes, performing slower and exaggerated movements when demonstrating novel actions to novices. The present study asked whether such modulations are also used to teach techniques of artistic expression. One domain where subtle movement modulations are crucial for achieving artistic expression is music. We investigated how skilled pianists modulate their playing to demonstrate to students the techniques required for conveying piece-related musical expressions, compared to performing the piece without didactic intentions. Expressions in the piece concerned either articulation (i.e., legato and staccato) or dynamics (i.e., forte and piano). The pianists played either with the goal to perform the piece to an audience or with the goal to teach the respective techniques to novices.When intending to teach articulation, skilled pianists produced more exaggerated staccato. When intending to teach dynamics, they created a larger contrast between forte and piano at structurally important points. We found consistent results with a simple musical scale (Experiment 1) and a more naturalistic piece of music (Experiment 2). These findings suggest that action modulations are performed not only to attract learners' attention but to highlight the most relevant aspects of the actions to be learnt.
}

@@

Importantly, in some domains teaching requires not only demonstrating \emph{what} to do, but \emph{how} to perform actions. In artistic contexts, novices need to learn how exactly to perform actions by relying on particular techniques (Sloboda, 2000). For example, when playing a piece of music on the piano, it is not sufficient to press the keys in the correct order, but learners need to be able to implement expressive techniques (Juslin, 2003). This raises the question of whether and how experts modulate their actions during didactic demonstration in artistic contexts, where expressivity is an integral part of what learners need to acquire. Studying action modulations that serve the teaching of artistic techniques is of broader theoretical importance: It can show whether action modulations can be used to selectively highlight subtle aspects of actions that define expert performance.

@@

If experts rely on generic action modulations, it can be expected that they will play more slowly during didactic demonstration, regardless of the kind of expressive techniques to be taught. To the extent that experts use action modulations to support their teaching of specific expressive techniques, they should produce longer legato and shorter staccato when teaching articulation whereas they should produce louder forte and softer piano when teaching dynamics. Furthermore, one could speculate that they might produce modulations specifically at structurally important points that best highlight the technique to be taught. Importantly, they should avoid modulating irrelevant properties of expression, e.g., the smoothness of the sound while teaching dynamics.

In Experiment 1, we employed a simple musical scale to examine whether and how skilled pianists vary their performance depending on which expressive techniques (i.e., either articulation or dynamics) they were teaching. Experiment 2 was conducted to replicate our findings from Experiment 1 with a more naturalistic piece of music.

@@

In order to compare the mean IOIs between the teaching and performing condition, we conducted a Wiscoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (\emph{p} \textless{} 0.001). The Wiscoxon Signed-rank test revealed that participants played more slowly in the teaching condition {[}\emph{Mdn} = 189.55 (ms), \emph{IQR} = 26.74{]} than in the performing condition {[}\emph{Mdn} = 185.24 (ms), \emph{IQR} = 19.32{]} while playing the piece with the notated articulation (\emph{p} = 0.021; \emph{Figure \ref{fig:ioi-1}}).

@@

A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) revealed that there was a significant main effect of Subcomponent-LS (\emph{F}(1,30) = 1158, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.94) and a significant interaction between Condition and Subcomponent-LS (\emph{F}(1,30) = 8.31, \emph{p} = 0.007, \(\eta_G^2\) = 0.011; \emph{Figure \ref{fig:kot-1}}). Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that participants produced shorter staccato in the teaching condition {[}\emph{M} = -134.03 (ms), \emph{SD} = 19.86{]} than in the performing condition {[}\emph{M} = -128.35 (ms), \emph{SD} = 20.80{]} (\emph{p} = 0.004) while there was no significant difference in legato between the teaching {[}\emph{M} = 15.91 (ms), \emph{SD} = 15.49{]} and performing condition{[}\emph{M} = 14.00 (ms), \emph{SD} = 17.01{]} (\emph{p} = 0.32).

@@

A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) showed that there was a significant interaction between Condition and Subcomponent-LS (\emph{F}(1,30) = 7.38, \emph{p} = 0.011, \(\eta_G^2\) = 0.004; \emph{Figure \ref{fig:vel-1}}). However, post-hoc comparisons based on the estimated marginal means with Tukey adjustment failed to find a significant difference between the teaching {[}Legato; \emph{M} = 70.34, \emph{SD} = 6.45, Staccato; \emph{M} = 72.01, \emph{SD} = 8.74{]} and performing condition {[}Legato; \emph{M} = 70.68, \emph{SD} = 5.84, Staccato; \emph{M} = 70.42, \emph{SD} = 7.96{]} for each subcomponent (Legato; \emph{p} = 0.70, Staccato; \emph{p} = 0.07).

@@

The Wiscoxon Signed-rank test revealed no significant difference between the teaching condition {[}\emph{Mdn} = 186.29 (ms), \emph{IQR} = 21.83{]} and the performing condition {[}\emph{Mdn} = 186.27 (ms), \emph{IQR} = 19.58{]} while playing the piece with the notated dynamics (\emph{p} = 0.111; \emph{Figure \ref{fig:ioi-1}}).

@@

A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,30) = 370, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.74) and a significant interaction between Condition and Subcomponent-FP (\emph{F}(1,30) = 6.80, \emph{p} = 0.014, \(\eta_G^2\) = 0.003; \emph{Figure \ref{fig:vel-1}}). Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte in the teaching condition {[}\emph{M} = 82.78, \emph{SD} = 8.30{]} than in the performing condition {[}\emph{M} = 81.66, \emph{SD} = 7.43{]} (\emph{p} = 0.021) while there was no significant difference in piano between the teaching {[}\emph{M} = 60.63, \emph{SD} = 5.07{]} and performing condition{[}\emph{M} = 60.99, \emph{SD} = 4.40{]} (\emph{p} = 0.45).

@@

Additionally, we conducted an exploratory analysis by focusing on specific points at which each subcomponent changed from one to the other (i.e., forte to piano; FtoP, piano to forte; PtoF). These points could be structurally important in order to make a contrast between forte and piano. We calculated a KV difference for each interval subtracting the KV value of the current note from that of the following note. A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (FtoP vs.~PtoF) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,30) = 224, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.83) and a significant interaction of Condition and Subcomponent-FP (\emph{F}(1,30) = 25.74, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.032; \emph{Figure \ref{fig:vel-diff}}). Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano and from piano to forte in the teaching condition {[}FtoP; \emph{M} = -14.06, \emph{SD} = 7.20, PtoF; \emph{M} = 24.26, \emph{SD} = 9.75{]} than in the performing condition {[}FtoP; \emph{M} = -11.69, \emph{SD} = 5.97, PtoF; \emph{M} = 20.88, \emph{SD} = 8.67{]} (FtoP; \emph{p} = 0.002, PtoF; \emph{p} \textless{} 0.001).

@@

A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) revealed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,30) = 377, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.31; \emph{Figure \ref{fig:kot-1}}), reflecting more key overlap for forte notes compared to piano notes. Neither the main effect of Condition (\emph{F}(1,30) = 0.42, \emph{p} = 0.520, \(\eta_G^2\) = 0.00) nor the interaction between Condition and Subcomponent-FP (\emph{F}(1,30) = 0.000, \emph{p} = 0.99, \(\eta_G^2\) = 0.000).

The findings from Experiment 1 indicated that skilled pianists were able to modify their performance for teaching purposes. For IOIs (tempo), we found a small but significant difference only when playing the piece with the notated articulation. This could be due to the leading metronome, which was given prior to each performance. As participants tried to follow the ideal tempo through their performances, it was possible that there was only a tiny difference between the two conditions. Another possibility could be that slower performance may be more or only beneficial to highlight the relation between two notes (i.e.~to what extent two notes overlap) when teaching articulation whereas such relation was irrelevant when teaching dynamics. With regard to KOT (articulation) and KV (dynamics), participants successfully highlight the relevant aspect of the technique. Specifically, they produced shorter staccato when teaching articulation and produced louder forte when teaching dynamics. Moreover, our exploratory analysis demonstrated that participants made a larger contrast in dynamics at transition points where one of two subcomponents moved to the other (i.e., forte to piano or piano to forte). These findings confirmed that participants modulated their performance in systematic and fine-grained ways in order to teach expressive techniques.

One of the limitations of Experiment 1 is that we employed a simple musical scale, which was good to gain an experimental control but not a naturalistic stimulus to implement artistic expressions. In Experiment 2, we chose one piece of music from actual piano pieces and modified it for the purpose of the experiment. We predict that participants will played more slowly regardless of the techniques to be taught or only when teaching articulation based on the finding from Experiment 1. Also, we expect that participants will be able to highlight only the relevant aspects of the techniques to be taught. More precisely, they should be able to produce longer legato and shorter staccato while teaching articulation. When teaching dynamics, they should be able to produce louder forte and softer piano. Given the findings we observed in Experiment 1, participants will make a larger contrast between forte and piano at transition points.

@@

@@

A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) showed that there was a significant main effect of Subcomponent-LS (\emph{F}(1,18) = 2324, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.98) and a significant interaction between Condition and Subcomponent-LS (\emph{F}(1,18) = 6.90, \emph{p} = 0.017, \(\eta_G^2\) = 0.01; \emph{Figure \ref{fig:kor-2}}). Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that participants produced shorter staccato in the teaching condition than in the performing condition (\emph{p} = 0.001) while there was no significant difference in legato between the teaching and performing condition (\emph{p} = 0.85).

@@

A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) showed that there was a significant main effect of Subcomponent-LS (\emph{F}(1,18) = 8.15, \emph{p} = 0.011, \(\eta_G^2\) = 0.0400; \emph{Figure \ref{fig:vel-2}}), reflecting louder sound for legato notes compared to staccato notes. Neither the main effect of Condition (\emph{F}(1,18) = 0.46, \emph{p} = 0.506, \(\eta_G^2\) = 0.0008) nor the interaction between Condition and Subcomponent-LS (\emph{F}(1,18) = 8.15, \emph{p} = 0.799, \(\eta_G^2\) = 0.0001).

@@

A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,18) = 126.10, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.7190) and a significant interaction between Condition and Subcomponent-FP (\emph{F}(1,18) = 6.11, \emph{p} = 0.024, \(\eta_G^2\) = 0.0217; \emph{Figure \ref{fig:vel-2}}). Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte and softer piano in the teaching condition than in the performing condition (Forte; \emph{p} = 0.036, Piano; \emph{p} = 0.028).

@@

A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (FtoP vs.~PtoF) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,18) = 110.64, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.8167) and a significant interaction of Condition and Subcomponent-FP (\emph{F}(1,18) = 9.77, \emph{p} = 0.006, \(\eta_G^2\) = 0.0641; \emph{Figure \ref{fig:vel-diff}}). Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano and from piano to forte in the teaching condition than in the performing condition (FtoP; \emph{p} = 0.017, PtoF; \emph{p} = 0.003).

@@

A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,18) = 559, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.93), reflecting more key overlap for forte notes compared to piano notes. Neither the main effect of Condition (\emph{F}(1,18) = 0.016, \emph{p} = 0.902, \(\eta_G^2\) = 0.00) nor the interaction between Condition and Subcomponent-FP (\emph{F}(1,18) = 0.576, \emph{p} = 0.458, \(\eta_G^2\) = 0.00; \emph{Figure \ref{fig:kor-2}}).

The results from Experiment 2 demonstrated that we successfully replicated our earlier findings that skilled pianists were able to modulate their performance for teaching purposes. As Experiment 1, we observed slower performance only while teaching the notated articulation. With regard to KOR (articulation), participants highlighted articulation by producing shorter staccato. Again, there was no significant difference in legato between the two conditions as Experiment 1. In term of KV (dynamics), participants produced louder forte and softer piano, which showed a clearer pattern of exaggeration compared to Experiment 1 where only softer piano was observed. Furthermore, we again found that participants made a larger contrast between forte and piano bidirectionally at transition points. To sum, Experiment 2 confirmed that participants modified their performance in systematic and fine-grained ways in order to teach expressive techniques with a naturalistic piece of music.

@@

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-kor-2-1} \caption{\label{fig:kor-2}Experiment 2: KOR when playing the piece with either articulation (left) or dynamics (right). Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box. Significance levels: * \textit{p} < .05, ** \textit{p} < .01, *** \textit{p} < .001}\label{fig:plot-kor-2}
\end{figure}

@@

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-vel-diff-1} \caption{\label{fig:vel-diff}Experiment 1, 2: KV Difference at transition points when playing the piece with dynamics (left; Experiment 1, right; Experiment 2). Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box. Significance levels: * \textit{p} < .05, ** \textit{p} < .01, *** \textit{p} < .001}\label{fig:plot-vel-diff}
\end{figure}

@@

The aim of the study was to investigate how expert pianists could modulate their performance in order to teach expressive artistic techniques. Overall, the findings of the two experiments showed that pianists performed one and the same piece differently depending on whether they played with the intention to teach expressive techniques or with the intention to perform the piece for an audience. When playing with the intention to teach, pianists selectively highlighted relevant aspects of artistic expressions. In particular, they exaggerated staccato when playing with the intention to teach expressions concerning articulation, and they exaggerated changes between forte and piano (in both directions) when playing with the intention to teach expressions concerning dynamics.

Although participants tended to play slower in the teaching condition than in the performing condition in general, we found a significant difference only for articulation, not for dynamics. As we mentioned in Experiment 1, it could be that the metronome beats given prior to each performance discouraged participants from deviating from the prescribed tempo. Another possibility is that expression is actually best taught when leaving the tempo unchanged when it is irrelevant. This would imply that general didactic modulations like slowing down, reported in research on parent-infant interaction, might be less useful in the context of teaching expressive artistic techniques where timing itself can be used to add an expression to a performance.

Participants exaggerated articulation for teaching purposes by highlighting staccato. The lack of a difference for legato might stem from a ceiling effect. When participants played the piece with dynamics, there was no difference between the two conditions with regard to KOT/KOR, which indicates that participants did not modulate irrelevant properties of the sound. However, participants made a larger contrast between forte and piano for teaching purposes to highlight the transition of dynamics. We expected there would be an overall exaggeration in KV while teaching dynamics and indeed found that participants produced louder forte and softer piano, but only in Experiment 2. The finding of making a larger contrast in dynamics at transition points suggests an effective way to highlight the technique as the loudness is determined relatively. Taken together, these findings demonstrate that expert pianists systematically modulated their sound depending on the specific technique they were trying to convey.

Musicians and music teachers tend to consider expressive skills as a performer's most important skills (Lindström, Juslin, Bresin, \& Williamon, 2003). In the present study, experts focused on the teaching of expressions that were notated. However, expression also has other facets (Juslin, 2003) that are not only piece-related but performer- or context-related. An interesting topic for future research is how experts teach not only expressive techniques as such but convey possibilities of interpretation.

One limitation of the study is that although we instructed participants to imagine a situation in which they were teaching musical expression to students, there was no feedback from actual students. Given that teachers modulate their demonstration throughout online interactions with learners (Okazaki, Muraoka, \& Osu, 2019), future studies are needed to investigate how experts dynamically adapt their performance to their student's skill level and demonstrated abilities.

The present findings extend earlier research on teaching-related action modulations. First, our study sheds light on the teaching of expressive techniques that are an integral part of skill acquisition in artistic domains. We showed that compared to an expressive performance baseline, experts made specific modulations in order to teach particular techniques. Second, the specificity of the observed modulations supports the idea that teaching comprises more than generic modulations like slowing down or overall exaggeration that may draw learners' attention. Rather, expert demonstrators seem to follow principles of relevance in communication (Sperber \& Wilson, 1995), highlighting only those aspects they are intending to demonstrate. How learners benefit from the perceptual and motor cues that come with specific exaggerations, and whether understanding the teacher's intentions explicitly adds to the learning success are important questions for future research.

@@

\leavevmode\hypertarget{ref-lindstrom_2003}{}%
Lindström, E., Juslin, P. N., Bresin, R., \& Williamon, A. (2003). ``Expressivity comes from within your soul'': A questionnaire study of music students' perspectives on expressivity. \emph{Research Studies in Music Education}, \emph{20}(1), 23--47. \url{https://doi.org/10.1177/1321103X030200010201}

@@

\leavevmode\hypertarget{ref-okazaki_2019}{}%
Okazaki, S., Muraoka, Y., \& Osu, R. (2019). Teacher-learner interaction quantifies scaffolding behaviour in imitation learning. \emph{Scientific Reports}, \emph{9}(1), 1--13. \url{https://doi.org/10.1038/s41598-019-44049-x}

@@

\leavevmode\hypertarget{ref-sperber_1995}{}%
Sperber, D., \& Wilson, D. (1995). \emph{Relevance: Communication and cognition, 2nd ed}. Malden: Blackwell Publishing.

@@

\endgroup

\hypertarget{supplementary}{%
\section{Supplementary Materials}\label{supplementary}}

\hypertarget{instructions}{%
\subsection{Instructions}\label{instructions}}

\hypertarget{experiment-1-1}{%
\subsubsection{Experiment 1}\label{experiment-1-1}}

In the teaching condition, participants were shown with the following instruction in front of the monitor (Italic sentences were highlighted with a yellow colour on a black background.); Now, play what you practised as if you were teaching it to students. Students already know how to produce the sequence of the tones and now are trying to \emph{learn how to perform the piece expressively} by listening to your performance. \emph{Do your best as a teacher to produce the piece according to the notation that you just practised}.

In the performing condition, participants were displayed with the following instruction in front of the monitor; Now, play what you practised as if you were performing it to an audience. \emph{Do your best as a performer to produce the piece according to the notation that you just practised}.

\hypertarget{experiment-2-1}{%
\subsubsection{Experiment 2}\label{experiment-2-1}}

In the teaching condition, we gave participants the exact same instruction as Experiment 1.

In the performing condition, participants were displayed with the following instruction in front of the monitor (Italic sentences were highlighted with a yellow colour on a black background.); Now, play what you practised as if you were performing it to an audience. \emph{Perform the piece expressively with your interpretation}. \emph{Do your best as a performer to produce the piece according to the notation that you just practised}.
